---
title: "Reproducible Research: Peer Assessment 1"
author: "Done by: DT"
date: "6/6/2020"
output: html_document
---

## Introduction

It is now possible to collect a large amount of data about personal
movement using activity monitoring devices such as a
[Fitbit](http://www.fitbit.com), [Nike
Fuelband](http://www.nike.com/us/en_us/c/nikeplus-fuelband), or
[Jawbone Up](https://jawbone.com/up). These type of devices are part of
the "quantified self" movement -- a group of enthusiasts who take
measurements about themselves regularly to improve their health, to
find patterns in their behavior, or because they are tech geeks. But
these data remain under-utilized both because the raw data are hard to
obtain and there is a lack of statistical methods and software for
processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring
device. This device collects data at 5 minute intervals through out the
day. The data consists of two months of data from an anonymous
individual collected during the months of October and November, 2012
and include the number of steps taken in 5 minute intervals each day.

## Data

The data for this assignment can be downloaded from the course web
site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as `NA`)

* **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

* **interval**: Identifier for the 5-minute interval in which
    measurement was taken




The dataset is stored in a comma-separated-value (CSV) file and there
are a total of 17,568 observations in this
dataset.


# Loading and preprocessing the data
Show any code that is needed to

1. Load the data (i.e. read.csv())

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r load data, echo=TRUE}

library(lubridate)

setwd("C:/Users/Osmiu/my_git_project/RepData_peerAssessment1")

activity <- read.csv("activity.csv", header=TRUE, na.strings = "NA")


#format date class
activity$date <- ymd(activity$date)


```

# Quick check

```{r summary, echo=TRUE}
summary(activity)
```


```{r str, echo=TRUE}
str(activity)
```



# 1. What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

#To remove the NAs and sum steps by date
```{r q1}
activity_steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
```




# 2. Make a histogram of the total number of steps taken each day

```{r plot histogram}
hist(activity_steps_day$steps, xlab = "Steps per Day", main = "Total number of steps taken per day", col = "Purple")
```


2. Calculate and report the mean and median total number of steps taken per day

```{r report mean and median total number of steps}
mean_steps <- mean(activity_steps_day$steps)
median_steps <- median(activity_steps_day$steps)

x <- format(mean_steps, digit=1)
y <- format(median_steps, digit=1)
```

Mean total number of steps per day: **`r x`**  
Median total number of steps per day: **`r y`**



# What is the average daily activity pattern?

# 1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
#set up data frame by using Aggregate function for mean of steps over all days, for each interval
activity_steps_mean_int <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)

#Plot graph 
plot(activity_steps_mean_int$interval, activity_steps_mean_int$steps, type = "l", col = "Red", xlab = "Intervals", ylab = "Total  mean steps per interval", main = "Average number of steps taken per interval (NA removed)")
```

# 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
#what is the maximum mean steps value? (maximum steps on one given interval)
max_mean_steps <-max(activity_steps_mean_int$steps)
#for which interval are the numbers of steps per interval at the highest?
max_mean_steps_interval <- activity_steps_mean_int$interval[which(activity_steps_mean_int$steps==max_mean_steps)]

```
The maximum number of steps for a 5 minutes interval is **`r max_mean_steps`**,  
which is at **`r max_mean_steps_interval`**  interval.


# Imputing missing values
## 1. Calculate total number of missing values in the dataset
```{r}
x <- sum(is.na(activity))
```
The total number of missing values in the dataset is **`r x`**


## 2. Devise a strategy for filling in all of the missing values in the dataset


```{r}
Replace_NA_activity <- activity
Replace_NA_activity$steps[is.na(Replace_NA_activity$steps)] <- mean(Replace_NA_activity$steps,na.rm=TRUE)
Replace_NA_activity$steps <- as.numeric(Replace_NA_activity$steps)
Replace_NA_activity$interval <- as.numeric(Replace_NA_activity$interval)

# this is to check if there are any 'NA'
colSums(is.na(Replace_NA_activity))
```

## 3. Create new dataset with the missing data filled in
 

```{r}
summary(Replace_NA_activity)
```

## 4. Make a histogram of the total number of steps taken each day 


```{r}
Replace_NA_activity2 <- aggregate(steps ~ date, data = Replace_NA_activity, FUN=sum)
```

```{r}
summary(Replace_NA_activity2)
```



```{r plot histogram with NA replaced}
hist(Replace_NA_activity2$steps, xlab = "Steps per Day", main = "Total number of steps taken per day after replacing Na", col = "Green")
```

```{r}
mean_steps_NA_replaced <- mean(Replace_NA_activity2$steps)
median_steps_NA_replaced <- median(Replace_NA_activity2$steps)

x <- format(mean_steps_NA_replaced, digit=1)
y <- format(median_steps_NA_replaced, digit=1)
```

Mean total number of steps taken per day after replacing NA = **`r x`**.

Median total number of steps taken per day after replacing NA = **`r y`**.


# Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.



```{r}
Replace_NA_activity$dayType <- ifelse(weekdays(as.Date(Replace_NA_activity$date)) == "Saturday" | weekdays(as.Date(Replace_NA_activity$date)) == "Sunday", "Weekend", "Weekday")
```

```{r}

Replace_NA_activity$dayType <- factor(Replace_NA_activity$dayType)
```



```{r}
str(Replace_NA_activity)
```

###2. Panel plot containing time series plot


```{r}
#Create new data frame using aggregate function with steps, interval and  daytype

table_steps_interval_dayType <- aggregate(steps ~ interval + dayType, data = Replace_NA_activity, FUN = mean)

library(ggplot2)

```



```{r}
#plot panel plot

plot <- ggplot(table_steps_interval_dayType, aes(interval, steps, dayType))
plot + geom_line(color = "green3") + facet_grid(dayType~.) + labs(x = "Intervals", y = "Mean Steps", title = "Mean steps by interval by dayType")
```




